Title Goes Here

### 2021-04-09 analysis/paper/shceof.Rmd
Agrega la barra de colores a todas las figuras
Closes #30

```{r define-plot_ceof}
plot_ceof <- function(data, n = 1:2, formula = lev  ~  cEOF) {

  gdata <- data[, cut(eof[[1]], n)$right, by = .(season)]

  var <- data[, cut(eof[[1]], n)$sdev, by = .(season)] %>%
    .[, setNames(paste0(cEOF, " (", scales::percent(r2), ")"),
                 cEOF)]

@@

    geom_contour_fill(aes(z = Re(hgt), fill = ..level..),
                      breaks = AnchorBreaks(0, 0.005, exclude = 0)) +
    geom_contour2(aes(z = Im(hgt),  linetype = factor(sign(..level..))),
                  breaks = AnchorBreaks(0, 0.005, exclude = 0)) +

    scale_fill_divergent_discretised(guide = "none") +

    geom_qmap(~ .x[lat <= -20]) +
    scale_x_longitude() +
    scale_y_latitude() +
    scale_linetype_manual(values = c("1" = 1, "-1" = 2), labels = c("1" = "+", "-1" = "-"),
                          guide = "none") +
    facet_grid(formula,
               labeller = labeller(lev = lab_lev,
                                   cEOF = var)) +
    coord_quickmap(ylim = c(NA, -20)) +
    theme(legend.title = element_blank())
}
```

```{r ceofs-1, fig.cap = "Spatial patterns of the leading 2 cEOFs of zonal anomalies of geopotential height at 50 hPa and 200 hPa for the SON trimester and the period 1979 -- 2019. Real part in shading, imaginary part in contours.", fig.width=width_column, fig.height=width_column}
ceof[season == "SON"] %>%
  plot_ceof() +
  coord_polar() +
  axis_labs_smol +
  tag_facets(tag = "cr")
```

@@

(ref:fake-eof-cap) Row a: The three components used to create a synthetic dataset of 41 years of geopotential height at 200 hPa. Row b: The three EOF computed from the synthetic dataset with the variance explained by each EOF in parenthesis. Units are arbitrary. Negative values in blue and positive values in red.

@@

```{r fake-eof, fig.cap = "(ref:fake-eof-cap)", fig.width=3.7, fig.height=3.2}
north_thumb <- function(sdev, n){
  sdev <- sdev^2
  delta <- sqrt(2 / n) * sdev

@@

  .[season(time) == "SON"] %>%
  .[, FitLm(hgt, Real, Imaginary), .(lon, lat)] %>%
  rm_intercept() %>%
  .[, .(lon, lat, eof = estimate, EOF = term)] %>%
  rbind(pattern[, .(lon, lat, eof, EOF = "SAM")]) %>%
  rbind(fake_eof$left, use.names = TRUE) %>% 
   .[lat <= -20] %>%
  .[, EOF := factor(EOF, c("SAM", "Imaginary", "Real", paste0("EOF", 1:3)),
                    ordered = TRUE)] %>% 
  .[EOF %in% c("EOF1", "EOF3"), eof := -eof] %>%

  periodic(lon = c(0, 360)) %>%
  ggplot(aes(lon, lat)) +

  geom_contour_fill(data = ~.x[EOF %in% c("SAM", "Imaginary", "Real")],
    aes(z = eof), global.breaks = FALSE, breaks = AnchorBreaks(0, exclude = 0)) +
  scale_fill_divergent(guide = "none") +
  ggnewscale::new_scale_fill() +
  geom_contour_fill(data = ~.x[!(EOF %in% c("SAM", "Imaginary", "Real"))],
    aes(z = eof), global.breaks = FALSE, breaks = AnchorBreaks(0, exclude = 0)) +

  geom_contour_tanaka(aes(z = eof), global.breaks = FALSE,  breaks = AnchorBreaks(0, exclude = 0)) +
  scale_fill_divergent(guide = "none") +
  geom_qmap(~.x[lat <= -20]) +
  scale_x_longitude() +
  scale_y_latitude() +
  coord_polar() +
  facet_wrap(~EOF, ncol = 3, labeller = labeller(EOF = var)) +
  axis_labs_smol +
  theme(axis.text = element_blank()) +
  tag_facets("rc")
```

>>>>

```{r ceofs-1, fig.cap = "Spatial patterns of the leading 2 cEOFs of zonal anomalies of geopotential height at 50 hPa and 200 hPa for the SON trimester and the period 1979 -- 2019. Real part in shading, imaginary part in contours. Arbitrary units.", fig.width=width_column, fig.height=width_column*1.2}

gdata <- ceof[, eof[[1]]$right, by = .(season)] %>% 
  copy() %>% 
  .[, hgt := hgt/sd(abs(hgt))]

var <- ceof[, eof[[1]]$sdev, by = .(season)] %>%
  .[, setNames(paste0(cEOF, " (", scales::percent(r2), ")"),
               cEOF)]

@@

  geom_contour_fill(aes(z = Re(hgt), fill = ..level..),
                    breaks = AnchorBreaks(0, 1, exclude = 0)) +
  geom_contour2(aes(z = Im(hgt),  linetype = factor(sign(..level..))),
                breaks = AnchorBreaks(0, 1, exclude = 0)) +

  scale_fill_divergent_discretised(guide = guide_colorsteps_bottom()) +

  geom_qmap(~ .x[lat <= -20]) +
  scale_x_longitude() +
  scale_y_latitude() +
  scale_linetype_manual(values = c("1" = 1, "-1" = 2), labels = c("1" = "+", "-1" = "-"),
                        guide = "none") +
  facet_grid( lev  ~  cEOF,
             labeller = labeller(lev = lab_lev,
                                 cEOF = var)) +
  coord_quickmap(ylim = c(NA, -20)) +
  theme(legend.title = element_blank()) +
  coord_polar() +
  axis_labs_smol +
  tag_facets(tag = "cr")

@@

(ref:fake-eof-cap) Row a: The three components used to create a synthetic dataset of 41 years of geopotential height at 200 hPa. Row b: The three EOF computed from the synthetic dataset with the variance explained by each EOF in parenthesis. Units are arbitrary.

@@

```{r fake-eof, fig.cap = "(ref:fake-eof-cap)", fig.width=3.7, fig.height=3.3}
north_thumb <- function(sdev, n){
  sdev <- sdev^2
  delta <- sqrt(2 / n) * sdev

@@

  .[season(time) == "SON"] %>%
  .[, FitLm(hgt, Real, Imaginary), .(lon, lat)] %>%
  rm_intercept() %>%
  .[, .(lon, lat, eof = estimate, EOF = term)] %>%
  rbind(pattern[, .(lon, lat, eof, EOF = "SAM")]) %>%
  rbind(fake_eof$left, use.names = TRUE) %>% 
   .[lat <= -20] %>%
  .[, EOF := factor(EOF, c("SAM", "Imaginary", "Real", paste0("EOF", 1:3)),
                    ordered = TRUE)] %>% 
  .[EOF %in% c("EOF1", "EOF3"), eof := -eof] %>% 
  .[!(EOF %in% c("SAM", "Imaginary", "Real")), eof := eof/sd(eof)] %>% 
  .[EOF %in% c("SAM", "Imaginary", "Real"), eof := eof/sd(eof)] %>%

  # geom_contour_tanaka has problem with the perfectly symemtric SAM
  # adding just a tiny bit of noise solves it. 
  .[EOF == "SAM", eof := eof + rnorm(.N, sd = 1e-5)] %>%

  periodic(lon = c(0, 360)) %>%
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = eof, fill = ..level..), global.breaks = TRUE, 
                    breaks = AnchorBreaks(0, 0.5, exclude = 0)) +

  geom_contour_tanaka(aes(z = eof), global.breaks = TRUE, 
                      breaks = AnchorBreaks(0,0.5,  exclude = 0)) +
  scale_fill_divergent_discretised(NULL, guide = guide_colorsteps_bottom()) +
  geom_qmap(~.x[lat <= -20]) +
  scale_x_longitude() +
  scale_y_latitude() +
  coord_polar() +
  facet_wrap(~EOF, ncol = 3, labeller = labeller(EOF = var)) +
  axis_labs_smol +
  theme(axis.text = element_blank()) +
  tag_facets("rc")
```
